Synthesis of Rules for Network Intrusion Detectors

### 2019-10-24 paper/paper.tex
pass on minimization procedure

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
\usepackage[caption=false, font=footnotesize]{subfig}
%%%%%%%%%%%%% code listing
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

\begin{document}
%
\title{Synthesis of Rules for Network Intrusion Detectors}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Alcantara\Comment{\inst{1}}\Comment{\orcidID{0000-1111-2222-3333}} \and
Guilherme Padilha\Comment{\inst{1}}\Comment{\orcidID{1111-2222-3333-4444}} \and
Marcelo d'Amorim\Comment{\orcidID{0000-0002-1323-8769}}}
%
\authorrunning{Alcantara et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Network Intrusion Detection Systems (\nids{}) are a popular mechanism
used by system administrators to defend against network attacks. These
systems monitor the network traffic and flag suspicious network
behavior. Signature-based \nids\ do that by checking the network
traffic against a pre-defined set of rules, which can become obsolete
as attackers learn new strategies to circumvent existing defenses.
This paper proposes \tname{}\footnote{Abbreviation of \textbf{Sy}nthesis of
  Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a technique that automatically
synthesizes \nids\ rules from positive and negative examples (\ie{},
malicious and benign traffic). \tname{} formulates synthesis as an
optimization problem whose candidate solutions are rules that
maximizes the captured the positive traffic and minimizes the captured
negative traffic. \tname{} bootstraps the search with candidate
solutions that use data from the payload of the malicious message and
produces optimal rule candidates on output. We evaluated \tname{} on a
diverse set of attacks. Results indicate that \Fix{...}

@@

Suricata supports a total of \totoptions\ distinct options covering
\numproto\ distinct protocols. A total of 9 of these options are for
documentation.
\begin{wrapfigure}[15]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}
  \scalebox{0.7}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={dns, ftp, http, icmp, ip, smb, smtp, ssh,
          tcp, tls, udp},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(dns,10) (ftp,11) (http,20) (icmp,18)(ip,16) (smb,13) (smtp,11) (ssh, 10) (tcp, 32) (tls,18) (udp, 18)};
%        \legend{relevant, irrelevant}
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-rules-protocol}Number of relevant
    options per protocol supported by Suricata.}
\end{wrapfigure}
\Comment{For instance, the purpose of rule option \CodeIn{msg} is to print on
the output an informative message indicating what kind of intrusion
has been observed.} Figure~\ref{fig:distribution-rules-protocol} shows
the distributions of these options per protocol. The sum of
the number of protocol options is higher than \totoptions\ as some
options is shared across different protocols.

@@

\Fix{--------------------------- estou aqui}

The rule reported by the first step satisfies the criterion to only
capture positive traffic, however it is ``too'' specific---it is
unable to capture different manifestations of the same attack as the
rule would unlikely match exactly the options we found for the input
message. Some options added to the rule during the first step are
irrelevant for determining the attack and should be removed.

The goal of this step is to create a \emph{minimal} rule that captures
\emph{only} the positive malicious traffic. A rule is minimal if
discarding any option results in capturing negative (\ie{}, benign)
traffic. Recall that the previous maximization step produces a rule
that captures the positive traffic and all options are satisifed by
the positive traffic. Consequently, any subset of these options also
captures the positive traffic. The minimization procedure we propose
in this step builds on that observation. It iteratively discards
options from the candidate rule until no other options can be
removed. The procedure uses negative traffic to guide the
search. Public databases of negative traffic with \Fix{thousands} of
messages exist \Fix{cite cite} and can be used to increase the
accuracy of the minimization.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%\usepackage[margin=1in]{geometry}% http://ctan.org/pkg/geometry
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
\usepackage{booktabs}
\usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
\usepackage[caption=false, font=footnotesize]{subfig}
%%%%%%%%%%%%% code listing
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

\begin{document}
%
\title{Synthesis of Rules for Network Intrusion Detectors}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Alcantara\MyComment{\inst{1}}\MyComment{\orcidID{0000-1111-2222-3333}} \and
Guilherme Padilha\MyComment{\inst{1}}\MyComment{\orcidID{1111-2222-3333-4444}} \and
Marcelo d'Amorim\MyComment{\orcidID{0000-0002-1323-8769}}}
%
\authorrunning{Alcantara et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Network Intrusion Detection Systems (\nids{}) are a popular mechanism
used by system administrators to defend against network attacks. These
systems monitor the network traffic and flag suspicious network
behavior. Signature-based \nids\ do that by checking the network
traffic against a pre-defined set of rules, which can become obsolete
as attackers learn new strategies to circumvent existing defenses.
This paper proposes \tname{}\footnote{Abbreviation of \textbf{Sy}nthesis of
  Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a technique that automatically
synthesizes \nids\ rules from positive and negative examples (\ie{},
malicious and benign traffic). \tname{} formulates synthesis as an
optimization problem whose candidate solutions are rules that
maximizes the captured the positive traffic and minimizes the captured
negative traffic. \tname{} bootstraps the search with candidate
solutions that use data from the payload of the malicious message and
produces optimal rule candidates on output. We evaluated \tname{} on a
diverse set of attacks. Results indicate that \Fix{...}

@@

Suricata supports a total of \totoptions\ distinct options covering
\numproto\ distinct protocols. A total of 9 of these options are for
documentation.
\begin{wrapfigure}[15]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}
  \scalebox{0.7}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={dns, ftp, http, icmp, ip, smb, smtp, ssh,
          tcp, tls, udp},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(dns,10) (ftp,11) (http,20) (icmp,18)(ip,16) (smb,13) (smtp,11) (ssh, 10) (tcp, 32) (tls,18) (udp, 18)};
%        \legend{relevant, irrelevant}
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-rules-protocol}Number of relevant
    options per protocol supported by Suricata.}
\end{wrapfigure}
\MyComment{For instance, the purpose of rule option \CodeIn{msg} is to print on
the output an informative message indicating what kind of intrusion
has been observed.} Figure~\ref{fig:distribution-rules-protocol} shows
the distributions of these options per protocol. The sum of
the number of protocol options is higher than \totoptions\ as some
options is shared across different protocols.

@@

The rule reported in the first step satisfies the criterion to only
capture positive traffic. Unfortunately, the rule produced in the
first step is overconstrained---it is unable to capture different
manifestations of the same attack as the rule matches all options that
\tname\ found for the input message.

%%  Some options added to the rule
%% during the first step are irrelevant for determining the attack and
%% should be removed to obtain more general rules.

\tname{} leverages public datasets\Fix{cite cite} of negative traffic
to guide a search for more general rules, hence addressing the
overffiting issue. The goal of this step is to create rules that
captures traffic similar to the input malicious traffic. To achieve
that goal, \tname{} \emph{minimizes} the rules produced in the
previous step, preserving the invariant of \emph{only} capturing the
positive malicious traffic. A rule is minimal if discarding any option
results in capturing negative (\ie{}, benign) traffic.  The
minimization procedure we propose in this step builds on the
observation that all options from the rule produced in the previous
step are satisifed by the positive traffic. Consequently, any subset
of these options also captures the positive traffic. The procedure
iteratively discards options from the candidate rule until it detects
that the rule is minimal, \ie{}, it finds that no other options can be
removed. The procedure uses negative traffic to guide the search.

\begin{algorithm}[t!]
\textbf{INPUT:} overconstrained rule $or$ \\
\textbf{OUTPUT:} Set of minimal rules $R$\\
\vspace{-3ex}
%\noindent\begin{minipage}{\textwidth}
\captionof{algorithm}{Minimization algorithm}\label{algo1}
\begin{algorithmic}[1]
%  \Procedure{MinRule}{$or$}\Comment{The g.c.d. of a and b}
    \State $r\gets a\bmod b$
    \While{$r\not=0$}\Comment{We have the answer if r is 0}
      \State $a\gets b$
      \State $b\gets r$
      \State $r\gets a\bmod b$
    \EndWhile
    \State \textbf{return} $b$\Comment{The gcd is b}
%  \EndProcedure
\end{algorithmic}
%\end{minipage}%
\end{algorithm}

%% Public databases of negative traffic with
%% \Fix{thousands} of messages exist  and can be used to
%% increase the accuracy of the minimization.
